### 2022-06-14 Rmarkdown/paper-si.Rmd
paper/data

---
title: "Supplementary information for: *Sodium silicate and hexametaphosphate promote the release of (oxyhydr)oxide nanoparticles from corroding iron*"
author: "Benjamin F. Trueman, Javier Locsin, Evelyne Dor&eacute;, Kalli Hood, and Graham A. Gagnon"
date: "Centre for Water Resources Studies,* Department of Civil & Resource Engineering, Dalhousie University, 1360 Barrington St., Halifax, Nova Scotia, Canada B3H 4R2  \n&nbsp;   \n&nbsp;*Corresponding author  \n&nbsp;E-mail: benjamin.trueman@dal.ca    \n&nbsp;Tel:902.494.6070  \n&nbsp;Fax: 902.494.3105"
output:
  word_document:
    keep_md: no
    reference_docx: StylesTemplate.docx
  html_document: default
bibliography: references.bib
csl: environmental-science-and-technology.csl
---

This document contains X tables and X figures.

```{r setup, include=FALSE}
here::i_am("Rmarkdown/paper-si.Rmd")
knitr::opts_chunk$set(
  echo = FALSE,
  dev = "png",
  dpi = 600,
  fig.width = 3.33,
  fig.height = 4
)
options(dplyr.summarise.inform = FALSE)
palette <- wesanderson::wes_palette("Zissou1", 6, type = "continuous")
```

```{r packages, message=FALSE}
library("tidyverse")
library("fffprocessr")
source(here::here("R/functions.R"))
```

```{r theme}
theme_set(
  theme_minimal() + 
    theme(
      legend.position = "bottom",
      legend.margin = margin(r = 1, unit = "cm")
    )
)
```

```{r load, message=FALSE, results="hide"}

c(
  combined = "data/fff_combined.csv",
  mals_data = "data/mals_combined.csv",
  calib_icp = "data/icp_calibration_files_clean.csv",
  direct_quant = "data/icp_direct_quant.csv",
  qc_recovery = "data/qc_recoveries.csv",
  image_j = "data/image_j_clean.csv",
  proteins = "data/proteins_clean.csv",
  pdat = "data/polyp_clean.csv"
) %>% 
  map(~ read_csv(here::here(.x))) %>% 
  list2env(envir = .GlobalEnv)

```

```{r plot-names}

param_names <- tibble::tribble(
   ~param,                       ~plot_param,
  "208Pb",       "Pb (&mu;g L<sup>-1</sup>)",
    "31P",        "P (&mu;g L<sup>-1</sup>)",
   "57Fe",       "Fe (&mu;g L<sup>-1</sup>)",
   "LS90", "90&deg; light scattering signal",
"UV254_2",                "UV<sub>254</sub>",
   "ls90",         "Radius of gyration (nm)"
)

sample_names <- tibble::tribble(
          ~sample,                                                   ~plot_sample,
        "100SiO2",      "Sodium silicate (100 mg SiO<sub>2</sub> L<sup>-1</sup>)",
            "3PP",             "Sodium hexametaphosphate (3 mg P L<sup>-1</sup>)",
 "FeOx-pHcontrol",                                                      "Control",
     "100SiO2+Pb", "Sodium silicate + Pb (100 mg SiO<sub>2</sub> L<sup>-1</sup>)",
         "3PP+Pb",        "Sodium hexametaphosphate + Pb (3 mg P L<sup>-1</sup>)",
        "HM_2ppm",             "Sodium hexametaphosphate (2 mg P L<sup>-1</sup>)",
  "TrimetaP_2ppm",              "Sodium trimetaphosphate (2 mg P L<sup>-1</sup>)"
   )

```

```{r fig-s1, fig.cap="**Figure S1:** Fractograms representing pure solutions of sodium hexameta- and trimeta-phosphate.", fig.height = 2.5, fig.width = 3.33}

pdat %>%
  left_join(sample_names, by = "sample") %>% 
  filter(time > 10, sample != "blank") %>% 
  ggplot(aes(time, conc, col = plot_sample, group = interaction(date, sample))) + 
  geom_line() + 
  scale_color_manual(values = palette[c(1,4)]) +
  labs(
    x = "Retention time (min.)",
    y = expression("[P] (µg L"^-1*")"),
    col = NULL
  ) + 
  guides(col = guide_legend(nrow = 2)) +
  theme(
    legend.text = ggtext::element_markdown()
  )

```

```{r checks, eval=FALSE}

# check that peak broadening is minimal between UV and ICP:

combined %>% 
  filter(
    param %in% c("UV254_2", "57Fe"),
    sample != "FeOx-pHcontrol",
    date < "2021-12-01"
  ) %>% 
  group_by(date, sample, param) %>% 
  mutate(conc = conc / max(conc)) %>% 
  ggplot(aes(time, conc, col =  param, group = interaction(date, param))) + 
  facet_wrap(vars(sample), scales = "free_y") +
  geom_line()

```

```{r mals-plot, eval=FALSE}

# Plot the data at all 21 angles

mals_data %>% 
  filter(time > 10, date < "2021-12-01") %>% 
  ggplot(aes(time, conc, col = sample, group = interaction(date, sample))) + 
  facet_wrap(
    vars(fct_reorder(param, str_extract(param, "\\d+") %>% as.numeric())), 
    scales = "free_y"
  ) + 
  geom_line() + 
  theme(legend.position = "bottom") +
  guides(col = guide_legend(nrow = 2)) +
  labs(col = NULL)
```