### 2022-07-07 README.Rmd

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Sodium silicate and hexametaphosphate promote the release of (oxyhydr)oxide nanoparticles from corroding iron

This repository contains the R code and other materials for the paper "Sodium silicate and hexametaphosphate promote the release of (oxyhydr)oxide nanoparticles from corroding iron" (soon to be available as a preprint). When rendering the .Rmd file, run `R/render_paper.R` to get the proper formatting for Arabic numeral citations.