### 2022-10-20 README.Rmd
update readme with link to published article

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Sodium silicate and hexametaphosphate promote the release of (oxyhydr)oxide nanoparticles from corroding iron

This repository contains the R code and other materials for the [paper](https://doi.org/10.1039/D2VA00199C) (and [preprint](https://doi.org/10.26434/chemrxiv-2022-rtzqm-v2)) "Sodium silicate and hexametaphosphate promote the release of (oxyhydr)oxide nanoparticles from corroding iron". When rendering the .Rmd file, run `R/render_paper.R` to get the proper formatting for Arabic numeral citations.